---
title: "IPL Analysis"
author: "Neilkesh Byrav H S"
date: "6 June 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries used
```{r Libraries, message=FALSE, warning=FALSE}
library(DataExplorer)
library(dplyr)
library(knitr)
library(ggplot2)
library(reshape2)
#Split the RT's
library(splitstackshape)
# Using gridExtra
library(gridExtra)
library(scales)
```

#Reading the files
```{r}
ipl = read.csv("C://Users//Administrator//Documents//Datasets//IPL//datasets//train//ball_by_ball_data.csv")
team_key  = read.csv("C://Users//Administrator//Documents//Datasets//IPL//datasets//train//key_teams.csv")
match_data  = read.csv("C://Users//Administrator//Documents//Datasets//IPL//datasets//train//match_data.csv")
player_rosters  = read.csv("C://Users//Administrator//Documents//Datasets//IPL//datasets//train//player_rosters.csv")
ipl_test = read.csv("C://Users//Administrator//Documents//Datasets//IPL//datasets//test//player_predictions.csv")
matches_2018 = read.csv("C://Users//Administrator//Documents//Datasets//IPL//datasets//test//matches_2018.csv")
```

#Exploratory Data Analysis
```{r warning=FALSE}
#Structure of IPL dataset
str(ipl)
#Plotting the structure
plot_str(ipl)


#Information about the dataset
introduce(ipl)

#Plotting the missing values
plot_missing(ipl) # Dataset does not contain any missing data at all. 
#The missing data column signifies that 95.06% were not outs

#Histograms
plot_histogram(ipl)
#The only interesting feature we get out of this is teh number of overs gradually decrease, as not all matches are completed with teams playing all 20 overs.

#Correlation
plot_correlation(ipl, maxcat = 5L, use = "pairwise.complete.obs")
#We can see that the entire data set is poorly correlated with each other, we can ignore the batsman and bowler id's, as they are the same.
```

#Data Transformation
###Adding Points,Type of match and Result per team
```{r warning=FALSE}
#Adding Points,Type of match and Result per team
match_data = cSplit(match_data, 'match_id','_', 'wide',drop = FALSE)
match_data$match_id_1 = NULL
colnames(match_data)[20]  = "match.no"
match_data_temp  = match_data %>% group_by(season) %>% summarise(Max_matches = max(match.no))
match_data = merge(x = match_data, y = match_data_temp, by = "season", all = TRUE)
match_data$match_type = ifelse(match_data$match.no<(match_data$Max_matches-3),"league","playoff")


match_data$Max_matches = NULL
match_data$team1_id = as.character(match_data$team1_id)
match_data$team2_id = as.character(match_data$team2_id)
match_data$winner = as.character(match_data$winner)
match_data$loser  = ifelse(match_data$result== "normal",
                ifelse(match_data$winner==match_data$team1_id,match_data$team2_id,match_data$team1_id),"")

match_data$team1_result = ifelse(match_data$result=="no result","no result",ifelse(match_data$result=="tie" ,"Tie",ifelse(match_data$winner==match_data$team1_id,"Win","Loss")))

match_data$team2_result = ifelse(match_data$result=="no result","no result",ifelse(match_data$result=="tie" ,"Tie",ifelse(match_data$winner==match_data$team2_id,"Win","Loss")))

match_data$team1_points = ifelse(match_data$match_type=="league",
                                 ifelse(match_data$team1_result== "Tie",1,                                                                                ifelse(match_data$team1_result== "no result",1,                                                                          ifelse(match_data$team1_result== "Win",2,0))),0)

match_data$team2_points = ifelse(match_data$match_type=="league",
                                 ifelse(match_data$team2_result== "Tie",1,                                                                                ifelse(match_data$team2_result== "no result",1,                                                                          ifelse(match_data$team2_result== "Win",2,0))),0)

#Splitting into two datframes for two teams
match_data$overal_match_no = seq.int(nrow(match_data)) 
match_data1  = match_data
match_data2 =  match_data
#Team1 Split
colnames(match_data1) [5] = "team"
colnames(match_data1) [23] = "team_result"
colnames(match_data1) [25] = "team_points"
match_data1$team2_id = NULL
match_data1$team2_result = NULL
match_data1$team2_points = NULL
#Team2 Split
colnames(match_data2) [6] = "team"
colnames(match_data2) [24] = "team_result"
colnames(match_data2) [26] = "team_points"
match_data2$team1_id = NULL
match_data2$team1_result = NULL
match_data2$team1_points = NULL

#binding both
match_data2 = rbind(match_data1,match_data2)
match_data2 = match_data2 %>% arrange(overal_match_no)
rm(match_data1,match_data_temp)

#Points per team 

points_data = match_data2[!match_data2$match_type=="playoff",] %>% group_by(season,team) %>% summarise(Points = sum(team_points),matches = max(n())) %>% mutate(matches,max_matches = max(matches)) %>% mutate(max_matches,max_points = max_matches*2) %>% mutate(Perc_points_scored  = round(Points/max_points,4)*100) %>% select("season","team","Perc_points_scored")
ggplot(points_data,aes(x = as.factor(season),y= Perc_points_scored)) + geom_line(size = 3)  + theme_bw() + labs( 
               y="Percentage of points scored", x="Seasons", title="Competition across season")

#The shorter the lines across seasons indicates tight competition during the legaue stages
```

#Team Stats across all seasons
```{r}
#Adding columns
match_data2$win_count  =ifelse(match_data2$team_result == "Win",1,0)
match_data2$loss_count  =ifelse(match_data2$team_result == "Loss",1,0)
match_data2$tie_count  =ifelse(match_data2$team_result == "Tie",1,0)
match_data2$nr_count  =ifelse(match_data2$team_result == "no result",1,0)
match_data2$count  = 1
match_data2$toss_winner = as.character(match_data2$toss_winner)
match_data2$toss_decision = as.character(match_data2$toss_decision)
match_data2$toss_win_count = ifelse(match_data2$team==match_data2$toss_winner,1,0)
match_data2$bat_first_count = ifelse(match_data2$toss_decision=="bat",1,0)
match_data2$field_first_count = ifelse(match_data2$toss_decision=="field",1,0)

#Team stats with points/wins/loss (Overall)

team_stats_2  = match_data2 %>% group_by(team) %>% summarise(Matches = sum(count),Wins = sum(win_count),Losses = sum(loss_count),Ties = sum(tie_count),No_result = sum(nr_count),Points = sum(team_points),Win_perc = round(100*sum(win_count)/sum(count),2),Toss_wins = sum(toss_win_count),Toss_win_perc = round(100*sum(toss_win_count)/sum(count),2) ,Bat_first = sum(bat_first_count),Field_first = sum(field_first_count))

#Team stats with points/wins/loss (SeasonWise)

x = unique(match_data2$season)
flag = x[1]
for (i in x) {
  y = subset(match_data2,season == i)
  
  b  = y %>% group_by(team) %>% summarise(Matches = sum(count),Wins = sum(win_count),Losses = sum(loss_count),Ties = sum(tie_count),No_result = sum(nr_count),Points = sum(team_points),Win_perc = 100*sum(win_count)/sum(count),Toss_wins = sum(toss_win_count),Toss_win_perc = 100*sum(toss_win_count)/sum(count) ,Bat_first = sum(bat_first_count),Field_first = sum(field_first_count))
  b$season = i

if(i==flag)
  {
  b2 = b
  }else{b2=rbind(b2,b)}

}
team_stat_seasonwise2 = b2
rm(b,b2,y,i,x,flag)

#Mutate new columns for analysis

ipl = ipl %>% mutate(Batsman_runs_count = if_else(batsman_runs == 0, 0, 1))
ipl$total_runs = ipl$batsman_runs + ipl$extra_runs
ipl = ipl %>% mutate(Total_runs_count = if_else(total_runs == 0, 0, 1))
ipl = ipl %>% mutate(Sixes = if_else(batsman_runs == 6, 1, 0))
ipl = ipl %>% mutate(Fours = if_else(batsman_runs == 4, 1, 0))
ipl = ipl %>% mutate(Wickets = if_else(dismissal_kind == '', 0, 1))
ipl = ipl %>% mutate(Dots = if_else(batsman_runs == 0, 1, 0))
ipl_def = ipl

#Team stats
#Number of matches
Team_matches = ipl %>% group_by(match_id,inning) %>% summarise(Teams = unique(batting_team))
Team_matches = subset(Team_matches,inning <3)
Team_matches$Count = 1
Team_matches = Team_matches %>% group_by(Teams) %>% summarise(Matches =sum(Count)) 

Teams_best1 = ipl %>% group_by(batting_team) %>% summarise(Runs = sum(total_runs),Sixes = sum(Sixes),Fours = sum(Fours),Strike_rate = sum(total_runs)/n()*100,Bat_avg = sum(total_runs)/sum(Wickets))
Teams_best2 = ipl %>% group_by(bowling_team) %>% summarise(Wickets = sum(Wickets),Bowl_avg = sum(total_runs)/sum(Wickets))
Teams_best = cbind(Teams_best1,Teams_best2)
Teams_best = cbind(Teams_best,Team_matches)
rm(Teams_best1,Teams_best2)
Teams_best$Boundaries = Teams_best$Fours + Teams_best$Sixes
Teams_best = Teams_best[,c(1,11,2,8,6,9,5,4,3)] %>% arrange(-Matches)
colnames(Teams_best) [1]  = "team"



#Team Stats by season

ipl_all = merge(x = ipl, y = match_data, by = "match_id", all = TRUE)
ipl_all = subset(ipl_all,inning <3)

x = unique(ipl_all$season)
flag = x[1]
for (i in x) {
  y = subset(ipl_all,season == i)
a = y %>% group_by(match_id,inning) %>% summarise(Teams = unique(batting_team))
a$Count = 1
a = a %>% group_by(Teams) %>% summarise(Matches =sum(Count)) 

Teams_best1 = y %>% group_by(batting_team) %>% summarise(Runs = sum(total_runs),Sixes = sum(Sixes),Fours = sum(Fours),Strike_rate = sum(total_runs)/n()*100,Bat_avg = sum(total_runs)/sum(Wickets))
Teams_best2 = y %>% group_by(bowling_team) %>% summarise(Wick = sum(Wickets),Bowl_avg = sum(total_runs)/sum(Wickets))
b = cbind(Teams_best1,Teams_best2)
b = cbind(b,a)
rm(Teams_best1,Teams_best2)
b$Boundaries = b$Fours + b$Sixes
b = b[,c(1,11,2,8,6,9,5,4,3)] %>% arrange(-Matches)
b$season = i

if(i==flag)
  {
  b2 = b
  }else{b2=rbind(b2,b)}

}
team_stat_seasonwise = b2
rm(b,b2,y,a,i,x,flag)
colnames(team_stat_seasonwise) [1]  = "team"

#Merging all the stats

Overall_Team_stats = merge(Teams_best,team_stats_2,"team",all = T)
Overall_Team_stats$Matches.y = NULL
Overall_Team_stats_seasonwise  = merge(team_stat_seasonwise,team_stat_seasonwise2,c("season","team"),all = T)
rm(team_stat_seasonwise,team_stat_seasonwise2,Teams_best,team_stats_2)

kable(Overall_Team_stats)
```

#Does the toss affect the outcome of the match ?
```{r}
toss_effect = select(Overall_Team_stats,c("team","Win_perc","Toss_win_perc"))
colnames(toss_effect)  = c("team","Match_Wins","Toss_Wins") 
toss_effect2 = melt(toss_effect,id.vars = "team")
ggplot(toss_effect2,aes(team,value)) + geom_bar(aes(fill = variable), position = "dodge", stat="identity",width = .5) +theme_bw() + theme(axis.text.x = element_text(angle = 90)) + labs(subtitle="Grouped Bar plot", 
               y="Percentage", x="Team", title="Match Wins vs Toss Wins") + theme(legend.title=element_blank()) + theme(axis.text.y=element_blank(), axis.title.y=element_blank(), axis.ticks.y = element_blank())

#The successful teams in the competition have not relied on the toss completely.
```
#Runs per over
```{r}
runs2 = ipl %>% group_by(over) %>% summarise(Total_runs = sum(batsman_runs),Overs = n()/6)
runs2$runs.per.over = runs2$Total_runs/runs2$Overs
runs2 = runs2 %>% arrange(runs.per.over)
runs2$over = as.factor(runs2$over)
ggplot(runs2,aes(x=over,y=runs.per.over)) + geom_bar(stat = 'identity',fill = 'red') 
#Runrate after the powerplay reduces by a huge margin
```

#Importance of all the balls in the over
```{r}
Overs = ipl %>% group_by(ball) %>% summarise(N.balls = n(),Total_runs = sum(total_runs),Total_sixes = sum(Sixes),Total_Fours = sum(Fours),Total_Wickets = sum(Wickets),Total_extras = sum(extra_runs))
Overs$Total_runs.per.ball = (Overs$Total_runs/Overs$N.balls)
Overs$Sixes.per.ball = (Overs$Total_sixes/Overs$N.balls)
Overs$Fours.per.ball = (Overs$Total_Fours/Overs$N.balls)
Overs$Wickets.per.ball = (Overs$Total_Wickets/Overs$N.balls)
Overs$Total_extras.per.ball = (Overs$Total_extras/Overs$N.balls)


Overs = subset(Overs,ball<7)
Overs$ball = as.factor(Overs$ball)
Overs2 = Overs[,-c(2:7)]
area.color <- c( NA, NA,NA, "withcolor", NA, NA)
a = ggplot(Overs,aes(x=ball,y=Total_runs.per.ball,fill=area.color)) + geom_bar(stat = 'identity') +  labs(subtitle="Total_Runs/Ball",y="Average Total Runs", x="Ball") +  theme_bw() + theme(legend.position="none") +  theme(axis.title.x=element_blank())

b = ggplot(Overs,aes(x=ball,y=Sixes.per.ball,fill=area.color)) + geom_bar(stat = 'identity') +  labs(subtitle="Sixes/Ball",y="Average Sixes", x="Ball") +  theme_bw() + theme(legend.position="none") + theme(axis.title.x=element_blank())

c = ggplot(Overs,aes(x=ball,y=Fours.per.ball,fill=area.color)) + geom_bar(stat = 'identity') +  labs(subtitle="Fours/Ball",y="Average Fours", x="Ball") +  theme_bw() + theme(legend.position="none") + theme(axis.title.x=element_blank())

d = ggplot(Overs,aes(x=ball,y=Wickets.per.ball,fill=area.color)) + geom_bar(stat = 'identity') +  labs(subtitle="Wickets/Ball",y="Average Wickets", x="Ball") +  theme_bw() + theme(legend.position="none") + theme(axis.title.x=element_blank())

e = ggplot(Overs,aes(x=ball,y=Total_extras.per.ball,fill=area.color)) + geom_bar(stat = 'identity') +  labs(subtitle="Total Extras/Ball",y="Average Total Extras", x="Ball") +  theme_bw() + theme(legend.position="none") + theme(axis.title.x=element_blank())

Plot_all = grid.arrange(a, b,c,d,e,nrow = 2, ncol = 3)
Plot_all
rm(a, b,c,d,e)
```
#Top 10 batsman in the IPL
```{r}
batsman_run = aggregate(batsman_runs~batsman,data=ipl,FUN = "sum")
top_batsman =head(batsman_run[order(batsman_run$batsman_runs,decreasing=TRUE),],10)


ggplot(top_batsman, aes(batsman, batsman_runs,fill=batsman)) + geom_col() + labs(subtitle="Bar Plot", 
               y="Total Runs", title="Top Batsman Runs In All Seasons") + theme(axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks.x = element_blank()) + geom_text(aes(label = batsman_runs), position = position_dodge(width = .9), vjust = -0.25)
```

#Merging the player information along with the batsman
```{r}
#Merging the player information
colnames(top_batsman)[1] = "Player"
player_rosters2 = player_rosters[player_rosters$Season=="2018",]
top_batsman = merge(top_batsman,player_rosters2,"Player")    
```

#Function to predict player score for the upcoming IPL season
```{r}
ipl_pred = function(files,toss_win,first_play)
{
  flag = files[1]
for (i in files) {

  player.id = i
df  = subset(ipl,batsman_id == player.id)

df  = df[,-c(6,7,8,9,10,11,14,15,16,17,18,19,20,21,22,23,24,25)]

df2 = df %>% group_by(match_id,bowler_id) %>% summarise(Runs = sum(batsman_runs),balls = n(),bat_team =unique(batting_team),bowl_team = unique(bowling_team)  )
df3 = df2
match_data_temp = select(match_data,c("match_id","venue_id","toss_winner","toss_decision")) 
df2 = merge(x = df2, y = match_data_temp, by = "match_id", all = TRUE)
df2 = na.omit(df2)
df2$toss_decision =as.character(df2$toss_decision)
df2$toss_win = ifelse(df2$bat_team==df2$toss_winner,1,0)

df2$first_play = ifelse(df2$toss_win==1,df2$toss_decision,ifelse(df2$toss_decision=="field","bat","field"))

df2$match_id = NULL
df2$toss_winner = NULL
df2$bat_team = NULL
df2$bowl_team = NULL
df2$toss_decision=NULL

df2$bowler_id = as.factor(df2$bowler_id)
df2$venue_id = as.factor(df2$venue_id)
df2$first_play = as.factor(df2$first_play)
df2$toss_win = as.factor(df2$toss_win)

Bowlers_avg = df3 %>% group_by(match_id,bowl_team) %>% summarise(Count = n()) 
Bowlers_avg2 = Bowlers_avg %>% group_by(bowl_team) %>% summarise(Avg = round(sum(Count)/n())) 
colnames(Bowlers_avg2)[1] = "team_id"

#Average Ball faced
Balls_avg = df2 %>% group_by(bowler_id) %>% summarise(balls = round(sum(balls)/n())) 

colnames(Balls_avg)[1] = "player_id"

#Test set preparartion

ipl_test2 = select(ipl_test,c("unique_id","match_id","player_id","team_id"))
ipl_test3 = merge(ipl_test2,matches_2018,by ="match_id",ALL=T)
player_match = subset(ipl_test3,player_id == player.id)

#Subsetting the matches played by player
ipl_test4 = ipl_test3[  ipl_test3$match_id %in% player_match$match_id, ]
ipl_test5 = ipl_test4[,-c(2,5,6,7,8)]
ipl_test6 = ipl_test5[! ipl_test5$team_id== player_match$team_id,]

ipl_test6 = ipl_test6[  ipl_test6$player_id %in% df2$bowler_id, ]
ipl_test6 = ipl_test6[  ipl_test6$venue_id %in% df2$venue_id, ]
ipl_test6$ID = seq.int(nrow(ipl_test6))
ipl_test6 = na.omit(merge(ipl_test6,Balls_avg,by = "player_id",all = T))
ipl_test6 = ipl_test6 %>% arrange(ID)

ipl_test6 = na.omit(merge(ipl_test6,Bowlers_avg2,by ="team_id",all=T))
ipl_test6 = ipl_test6 %>% arrange(ID)
ipl_test6$ID = NULL

#Selecting random bowlers
for (j in 1:50) {

#set.seed(101)
ipl_test7 = ipl_test6 %>% group_by(match_id) %>% mutate(id = row_number(sample(25, size = n(), replace = TRUE)))
ipl_test7$flag = ifelse(ipl_test7$id >ipl_test7$Avg,1,0)
ipl_test7 = ipl_test7[!ipl_test7$flag==1,]
final_bind = ipl_test7
final_bind = final_bind[,-c(2,4,5,6,7,8)]
ipl_test7 = ipl_test7[,-c(1,3,6,7,8)]
colnames(ipl_test7)[1] = "bowler_id"

ipl_test7$bowler_id = as.factor(ipl_test7$bowler_id)
ipl_test7$venue_id = as.factor(ipl_test7$venue_id)
ipl_test7$toss_win = toss_win
ipl_test7$first_play = first_play

ipl_test7$toss_win = as.factor(ipl_test7$toss_win)
ipl_test7$first_play = as.factor(ipl_test7$first_play)

train = df2
test = ipl_test7

#Model
runs_lm2 = lm(Runs ~ .,train)
run_lm_score2 = predict(runs_lm2, test)

final = cbind(ipl_test7,round(run_lm_score2))
colnames(final)[6] = "runs"
final.bind =as.data.frame(final_bind)
final2 = cbind(final.bind,final)

score_per_match = final2 %>% group_by(match_id) %>% summarise(Runs = sum(runs),Balls = sum(balls))
total_score = as.data.frame(sum(score_per_match$Runs))
colnames(total_score) = "runs"

if(j==1)
  {
    total_scores = total_score
}else{total_scores=rbind(total_scores,total_score)}

}

Average_score = as.data.frame(round(mean(total_scores$runs)))
colnames(Average_score)= "Total_Runs"
Average_score$player_id = player.id
Predicted_score = merge(Average_score,player_rosters2,by.x = "player_id" )

if(i==flag)
  {
    Top_Ten_Players = Predicted_score
}else{Top_Ten_Players=rbind(Top_Ten_Players,Predicted_score)}

}
  
  return(Top_Ten_Players)
}  
```

#Predicted scores for the upcoming season
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
files = top_batsman$player_id

#Bat first and Toss win
Top_Ten_Players = ipl_pred(files,1,"bat")
Top_Ten_Players = Top_Ten_Players %>% arrange(-Total_Runs)
```

#Plotting the predicted scores
```{r}
ggplot(Top_Ten_Players, aes(Player, Total_Runs,fill=Player)) + geom_col() + labs(subtitle="Bar Plot", 
               y="Total Runs", title="Predicted total scores for 2018 IPL Season") + theme(axis.text.x=element_blank(), axis.title.x=element_blank(), axis.ticks.x = element_blank()) + geom_text(aes(label = Total_Runs), position = position_dodge(width = .9), vjust = -0.25)
```

#Comparing the scores with the actual season scores
```{r}
#Considering players who played all 14 matches
players_play_all_games = c("Virat Kohli","AB De Villiers","Suresh Raina","Robin Uthappa","Rohit Sharma","Shikhar Dhawan","Mahendra Singh Dhoni")
#Filter
Top_Players =  Top_Ten_Players[Top_Ten_Players$Player %in% players_play_all_games,]
#Manually Entering the total scores from "www.iplt20.com"
#All the players have played all the matches upto the league stage(scores are counted only upto the league stage)
Top_Players$Actual_runs = c("530","480","391","346","286","437","446")

Top_Players2 = select(Top_Players,"Player","Actual_runs","Total_Runs")
colnames(Top_Players2) = c("Player","Actual_Runs","Predicted_Runs")
Top_Players2$Player = c("Virat","AB De","Raina","Uthappa","Rohit","Dhawan","Dhoni")
Top_Players2 = melt(Top_Players2,id.vars = "Player")

ggplot(Top_Players2,aes(Player,value)) + geom_bar(aes(fill = variable), position = "dodge", stat="identity",width = .5) +theme_bw() + theme(axis.text.x = element_text(angle = 90)) + labs(subtitle="Grouped Bar plot", 
               y="Runs", x="Players", title="Actual Score vs Predicted Score") + theme(legend.title=element_blank())  
```


#Assesing the Linear Regresson Variables
```{r}
#Let us take Virat Kohli as Example
player.id = 149
df  = subset(ipl,batsman_id == player.id)

df  = df[,-c(6,7,8,9,10,11,14,15,16,17,18,19,20,21,22,23,24,25)]

df2 = df %>% group_by(match_id,bowler_id) %>% summarise(Runs = sum(batsman_runs),balls = n(),bat_team =unique(batting_team),bowl_team = unique(bowling_team)  )
df3 = df2
match_data_temp = select(match_data,c("match_id","venue_id","toss_winner","toss_decision")) 
df2 = merge(x = df2, y = match_data_temp, by = "match_id", all = TRUE)
df2 = na.omit(df2)
df2$toss_decision =as.character(df2$toss_decision)
df2$toss_win = ifelse(df2$bat_team==df2$toss_winner,1,0)

df2$first_play = ifelse(df2$toss_win==1,df2$toss_decision,ifelse(df2$toss_decision=="field","bat","field"))

df2$match_id = NULL
df2$toss_winner = NULL
df2$bat_team = NULL
df2$bowl_team = NULL
df2$toss_decision=NULL

df2$bowler_id = as.factor(df2$bowler_id)
df2$venue_id = as.factor(df2$venue_id)
df2$first_play = as.factor(df2$first_play)
df2$toss_win = as.factor(df2$toss_win)

Bowlers_avg = df3 %>% group_by(match_id,bowl_team) %>% summarise(Count = n()) 
Bowlers_avg2 = Bowlers_avg %>% group_by(bowl_team) %>% summarise(Avg = round(sum(Count)/n())) 
colnames(Bowlers_avg2)[1] = "team_id"

#Average Ball faced
Balls_avg = df2 %>% group_by(bowler_id) %>% summarise(balls = round(sum(balls)/n())) 

colnames(Balls_avg)[1] = "player_id"

train = df2

#Model
runs_lm = lm(Runs ~ .,train)
```

#Testing across the same dataset
```{r}
set.seed(100)
indexes = sample(1:nrow(df2), size=0.7*nrow(df2))
train = df2[indexes,]
test= df2[-indexes,]
test2 = test[  test$bowler_id %in% train$bowler_id, ]
test2 = test2[  test2$venue_id %in% train$venue_id, ]

runs_lm3 = lm(Runs ~ ., train)
summary(runs_lm3)
act = test2$Runs
test2$Runs =NULL

run_lm_score = predict(runs_lm3, test2)

#Get a vector of the differences between the predictions and actual results, and get its mean and standard deviation.
act_pre = as.data.frame(act)
act_pre$pre = round(run_lm_score)
act_pre$player_id  =test2$bowler_id
player_rosters3 = player_rosters %>% group_by(player_id) %>% summarise(Player = unique(Player))
act_pre = merge(act_pre,player_rosters3,by="player_id")
act_pre$pre = ifelse(act_pre$pre <0,0,act_pre$pre)
act_pre$balls = test2$balls 

Kohli_Perfor =   act_pre %>% group_by(Player) %>% summarise(Actual_Score = sum(act),Predicted_Score = sum(pre),Balls_Faced = sum(balls)) %>% arrange(-Actual_Score) %>% head(10)

Kohli_Perfor$Actual_Score_Rate  =  round(100*Kohli_Perfor$Actual_Score/Kohli_Perfor$Balls_Faced,2)
Kohli_Perfor$Predicted_Score_Rate = round(100*Kohli_Perfor$Predicted_Score/Kohli_Perfor$Balls_Faced,2)
Kohli_Perfor = Kohli_Perfor[,-c(2,3,4)]

Kohli_Perfor = melt(Kohli_Perfor,id.vars = "Player")

ggplot(Kohli_Perfor,aes(Player,value)) + geom_bar(aes(fill = variable), position = "dodge", stat="identity",width = .5) +theme_bw() + theme(axis.text.x = element_text(angle = 90)) + labs(subtitle="Actual Strike rate vs Predicted strike rate", y="Runs", x="Players", title="Kohli's Performance against Bowlers") + theme(legend.title=element_blank())  

diff = act_pre$act - act_pre$pre
mean(diff, na.rm = TRUE)
sd(diff, na.rm = TRUE)
```



